---
title: "class8"
author: "Miso Lee"
date: "April 26, 2018"
output: 
  html_document: 
    keep_md: yes
---

K means example-----------


Generate some example data for clustering
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```


use the kmeans() functions setting K to 2 and nstart=20

```{r}
km <- kmeans(x, centers = 2, nstart=20 )
```

inspect/print the results
```{r}
km
```

Q. How many points are in each cluster?
Q. What 'component' of your result object details: cluster size, assignement/membership, centers

```{r}
km$cluster
km$centers
km$size
```


plot x colored by the kmeans cluster assignment and add cluster centers as blue points

```{r}
plot(x, col=km$cluster)
points(km$centers, col = "blue", pch=15)
```


Q. Repeat for k=3, which one has the better total SS?

```{r}
km3 <- kmeans(x, centers = 3, nstart = 20)
km3
km3$withinss
```

########Hierarchial clustering example-----------

First we need to calculate point (dis)similarity

as the Euclidean distance between observations

```{r}
dist_matrix <- dist(x)
dist_matrix
```

lets look at the class

```{r}
class(dist_matrix)
View(as.matrix(dist_matrix))
dim(dist_matrix)
dim(as.matrix(dist_matrix))
```

The hclust() function returns a hierarchical
clustering model

```{r}
hc <- hclust(d = dist_matrix)
```

the print method is not so useful here

```{r}
hc
```

Create hierarchical cluster model: hc

```{r}
hc <- hclust(dist(x))
```

We can plot the results as a dendrogram
```{r}
plot(hc)
```


```{r}
# The hclust() function returns a hierarchical
# clustering model

hc <- hclust(d = dist_matrix)

```

Lets draw the tree

```{r}
plot(hc)
abline(h=6, col="red")
```

I can 'cut' the tree at any height to give our clusters...

```{r}
# Cut at height 6
cutree(hc, h=6)
```

cut into k grps

```{r}
cutree(hc, k=2)
```

apply()
```{r}
apply(x, 2, sd)
```




**YOUR TURN**

```{r}
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```

```{r}
plot(x)
```

```{r}
col <- as.factor( rep(c("c1", "c2", "c3"), each=50))
```

```{r}
plot(x, col=col)
```

Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters 

```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
```

```{r}
# To return 2 clusters I could cut at h=2.5 or set k=2
members2 <- cutree(hc, k=2)

# To return 3 clusters I could cut at h=2 or set k=3
members3 <- cutree(hc, k=3)

# Let's look at both
members <- cbind(k2=members2, k3=members3)
```

Plot

```{r}
plot(x, col=members3, pch=15)
```


**How to do PCA in R**

```{r}
# Generating example data to work with
mydata <- matrix(nrow=100, ncol=10)

# Let's label the rows  gene1, gene2 etc. to gene100
rownames(mydata) <- paste("gene", 1:100, sep="")

# Let's label the first 5 columns wt1, wt2, wt3, wt4 and wt5 and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c(paste("wt", 1:5, sep=""),
                      paste("ko", 1:5, sep=""))

# Fill in some fake read counts
for(i in 1:nrow(mydata)) {
  wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
  ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))
  mydata[i,] <- c(wt.values, ko.values)
}

head(t(mydata))
```

```{r}
# Let's do PCA
pca <- prcomp(t(mydata), scale=TRUE)
pca

# See what is returned by the prcomp() function
attributes(pca)

# A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2])

# Variance capture per PC
# How well are our PC's capturing our data spread (i.e. variance)
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

# Plot our screen plot
barplot(pca.var.per, main="Screen Plot", xlab="Principal Component", ylab="Percent Variation")
```

Color Vector
```{r}
# A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"

# Plot
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1(", pca.var.per[1], "%)"),
     ylab=paste0("PC2(", pca.var.per[2], "%)"))

# Click to identify which sample is which
identify(pca$x[,1], pca$x[,2],  labels=colnames(mydata))

##한번에 console에서 run하기
```


**UK FOODS**
```{r}
#reading UK FOODS
UK_foods <- read.csv("UK_foods.csv", row.names = 1)
View(UK_foods)
```







