---
title: "class 16"
author: "Miso Lee"
date: "May 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# installing package
# install.packages("tidyverse")

# load packages
library(tidyr)
library(dplyr)
library(ggplot2)
library(readr)

# loading file
exp_data <- read.csv("data/Experimental_results.csv")
head(exp_data)

# remove column "X" from a dataframe:
# exp_data <- exp_data[,-1]
exp_data <- select(exp_data, -X)
head(exp_data)
```

```{r}
# select function practice
Mut_str_env <- select(exp_data, Mut_ID, Strain, Environment)
head(Mut_str_env)
```

To make a graph, we need to give a function 2 variables (2 columns) to plot them against each other.

What variables are we going to use?

Lets rearrange our table to be able to plot the data easily. Instead on keeping information about barcode frequency in rows, we are going to create a column “Time” with time points and a column “Frequency” with corresponding barcode frequencies.

```{r}
# Information from a ggplot description page:
# ggplot(df, aes(x, y, <other aesthetics>))

exp_rearranged <- gather(exp_data, Generation, Frequency, H_0:H_96)
head(exp_rearranged)
```

You might have noticed that “Generation” column contains both “H” that stands for “hours” and numbers. Lets remove “H_” part from this column.

Check the syntax of “separate” function.

```{r}
# Separate values in "Generation" column into 2 columns
table_for_graph <- separate(exp_rearranged, Generation, into = c("H","Time"))
head(table_for_graph)
```

```{r}
# Remove column "H" using function "select"
table_for_graph <- select(table_for_graph, -H)
table_for_graph
```

You might have noticed that our table contains a lot of “NA” values. Go ahead and remove them with na.omit function. Don’t forget to check it’s syntax first!

```{r}
table_cleaned <- na.omit(table_for_graph)
table_cleaned$Time <- as.numeric(table_cleaned$Time)
head(table_cleaned)
```

Now the table is ready. How are we going to plot all the values? Do we need to separate them in any way? If yes, then how?

```{r}
# We might need to plot data for each strain separately..
DivAnc <- filter(table_cleaned, table_cleaned$Strain == "DivAnc")
L013 <- filter(table_cleaned, table_cleaned$Strain == "L013")

# make a plot for DivAnc strain
DivAnc_plot = ggplot(DivAnc) + geom_line(aes(x = Time, y = Frequency, group = BC_ID), alpha = .2, colour = "#000033") + ggtitle("DivAnc_SC3") + theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
DivAnc_plot

# make a plot for L013 strain
L013_plot = ggplot(L013) + geom_line(aes(x = Time, y = Frequency, group = BC_ID), alpha = .2, colour = "#CC6633") + ggtitle("L013_SC3") + theme(plot.title = element_text(hjust = 0.5)) + xlab("Time, hours") + ylab("Log10(Barcode frequency)")
L013_plot
```

Can we make 2 graphs at the same time?

```{r}
ggplot(table_cleaned) + geom_line(aes(x = Time, y = Frequency, group = BC_ID), alpha = .2, colour = "#000033") + facet_grid(.~Strain) + ggtitle("Barcode trajectories") + theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
```

Lets pick one mutation and check how it behaves in different strains

```{r}
# I've chosen Mut_ID==34
mut34 <- filter(table_cleaned, table_cleaned$Mut_ID == "34")
mut34  
```

```{r}
ggplot(mut34, aes(Time, Frequency, group = BC_ID,  color = BC_ID)) + geom_line() + theme(legend.position = "none") + facet_grid(.~Strain) + ggtitle("Mutation_34") + xlab("Time, hours") + ylab("Log10(Barcode frequency)") + theme(plot.title = element_text(hjust = 0.5))
```

Why do we have 2 clusters of barcodes?

Lets filter out barcodes with frequency > (-5) and use them for subsequent analysis.

```{r}
mut34_f <- filter(mut34, mut34$Frequency > (-5))
mut34_f
```

Plot again the same type of graph, but use filtered data. Make sure that you have done everything right.

```{r}
ggplot(mut34_f, aes(Time, Frequency, group = BC_ID, color = BC_ID)) + geom_line() + theme(legend.position = "none") + facet_grid(.~Strain) + ggtitle("Mutation_34") + xlab("Time, hours") + ylab("Log10(Barcode frequency)") + theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(mut34_f, aes(Time, Frequency, colour = BC_ID, group = BC_ID)) + geom_point() + geom_smooth(se = FALSE, method = "lm") + facet_grid(.~Strain) + theme(legend.position = "none") + ggtitle(paste("Mutation", 34, sep = "_")) + xlab("Time, hours") + ylab("Log10(Barcode frequency)")
```

Now you can choose a different mutation and check how it behaves across strains.

Now it’s time to estimate slope for each barcode. Lets greate a file that will contain information about BC_ID, Mut_ID, Strain, and estimated slope.

```{r}
# Lets become familiar with lm function:

# For this exercise, take the filtered data for mutation 34 (mut34_f) and filter out information about one barcode you like

# I have chosen BC_ID=25361 in DivAnc strain
BC_25361 <- filter(mut34_f, mut34_f$BC_ID == "25361", mut34_f$Strain == "DivAnc")
BC_25361
```

```{r}
#Lets plot frequency of this barcode:
BC_plot <- ggplot(BC_25361, aes(Time, Frequency, colour = BC_ID)) + geom_point() + theme(legend.position = "none") + ggtitle("BC_25361") + xlab("Time, hours") + ylab("Log10(Frequency)")
```

```{r}
#Lets use lm function to fit the line to these points:
# ggplot(BC_25361, aes(Time, Frequency, colour = BC_ID)) + geom_point() + geom_smooth(se = FALSE, method = "lm") + theme(legend.position = "none") + ggtitle("BC_25361") + xlab("Time, hours") + ylab("Log10(Frequency)")

BC_plot_lm <- BC_plot + geom_smooth(se = FALSE, method = "lm")
BC_plot_lm
```

```{r}
# Lets check what data does lm function return:
regression_model <- lm(Frequency~Time, BC_25361)
summary_data <- summary(regression_model)
summary_data
```

```{r}
# The information we are interested in is the value of Slope and Intercept of this line:
# Let's try to access them:

# Time
Time<-summary_data$coefficients[2]
Time
```

```{r}
# Intercept:
Intercept<-summary_data$coefficients[1]
Intercept
```

